The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-11-16 theoretical-consensus/article/article.tex
add appendix

>>>>

\section[Appendix. In-browser Proof-of-Work]{Appendix\\ {In-browser Proof-of-Work}}

Along with this research, a simple practical prototype was developed
to demonstrate the effectiveness of Proof-of-Work as a permissionless consensus protocol.
The prototype was developed with the idea of being accessible to any device that supports
the modern features of web browsers. Indeed, this prototype can run in a seemingly decentralized fashion,
being delivered as a static web page, without the inherent need for a centralized hosting or backend server.
The next sections will give more details about the implementation and design considerations, 
as well as a short set of remarks and future work.

\subsection{Implementation}

This prototype was developed and shipped as a static web page.
The base layer tooling choice resided in the \emph{Nuxt.js} framework, a meta-framework
for building web applications based on \emph{Vue.js}. With that established,
it became clear the need for a protocol for syncing and communicating in a 
peer-to-peer way between the devices. The choice resided in the \emph{GUN} ecosystem,
which ships a small but powerful set of tools for in-browser data synchronization.
Since there was \emph{TypeScript} support from all the major dependencies of this project,
this was the language chosen for the interactive web development. The cryptographic standards used in the 
\emph{PoW} algorithm were imported from the \emph{crypto-js} library. And, finally, HTML and CSS, with the help
of the \emph{Tailwind CSS} framework were used for styling and presentation.

A major requirement for such a system is the capacity for parallel computations, 
in order to be able to, simultaneously, listen for and validate proposed blocks, 
present the current results in the screen, and mine new blocks. The traditional 
\emph{JavaScript} computation model is a single-threaded environment that, naturally,
does not allow for the efficient and effective handling of all these scenarios, happening
at the same time. Most of the modern web browsers have a living solution for this problem, 
and that is based on \emph{Web Workers}. This \emph{API} was used in order to offload
the expensive mining computations and take that burden out of the main thread, leaving it
with the responsibility of listening for information updates, either from other peers, with
proposed blocks, or from the \emph{miner} threads, with new mined blocks, and decide and update
the user interface accordingly.

The remaining piece of the puzzle is the device discovery issue. The way the internet works does not
allow for direct device pairing and easy peer discovery. The modern web generally follows a 
client-server model with the centralization of resources. And, for this project, there was the need
for a mechanism that would allow the peers to connect to each other to then proceed to synchronize
the information. The \emph{GUN} ecosystem established a commonly accepted solution for this issue, that, 
very shortly, depends on setting up \emph{relay} peers that have somewhat static addresses and that can 
serve as reference points and pairing gateways for the nodes to discover and establish direct communications 
with their peers. For development and demonstration purposes, multiple relay nodes were set up, running 
as \emph{Docker} containers, with static ports and addresses, that could then be referenced in the web 
page generation, for the other mobile and dynamic peers to connect to and make contacts with the rest of the network.
The communication fashion is abstracted by the libraries used, but, generally, it happens either via \emph{WebSockets}
or \emph{WebRTC} technology. Figure \ref{fig:in-browser-pow-topology} shows the topology of the whole system.

\begin{figure*}[h]
  \centering
  \captionsetup{justification=centering}
  % \includegraphics{in-browser-pow-topology}
  \caption{blblb}
  \label{fig:in-browser-pow-topology}
\end{figure*}

\subsection{Remarks}

The prototype was tested, is demonstrable and works as intended. Some \emph{Nakamoto} chain-based
features were simplified, for the sake of maintaining a didactic complexity. For example, 
the Longest-Chain rule, in this prototype, works by having a temporary set of active blocks 
that get appended to the main chain if there are new blocks mined on top
of them. This setting enforces higher synchrony between the peers, as there is no mechanism for
general chain data synchronization. This also means that a new peer joining the network does not
need or does not ask for the rest of the chain history to start mining, but just the most recent block.
Again, this is justified by the purpose of this prototype which was to demonstrate the achievement
of consensus in a permissionless environment.

No extensive benchmarks or analysis were made, other than live tests with multiple
and heterogeneous devices. The basic tests involved one, two or more browser tabs
opened in the same device, competing for mining the blocks. The distribution of blocks mined
in a fixed and considerable interval was observed to be uniform for the participants. Another test was
conducted with a laptop and a mobile device, for which, for obvious reasons, the laptop
had a generally bigger advantage in the mining process.

\subsection{Future Work}
